Sampling deaths and OLS

### 2018-01-20 ols.Rmd
revision

---
title: "Sampling deaths and OLS"
author: "Göran Broström"
date: "January 19, 2018"
output: html_document
---

## Introduction

@@

## Data

@@

```{r readalldata}
source("R/get.file.R")
skume <- get.file("skume")
## Not needed(?):
##vb0Tab <- skume
##vb0Tab$civst <- relevel(vb0Tab$civst, ref = "married")
##vb0Tab <- aggregate(vb0Tab[, c("event", "exposure")],
##                    vb0Tab[, c("sex", "age", "period", "cohort", "civst", "hisclass", "urban")],
##                    FUN = sum)
## End: Not needed(?)
ols <- obs[, c("id", "sex", "birthdate", "region", "civst", "hisclass", "sluttyp", "enter", "exit")]
ols$event <- ols$sluttyp == 2
ols$sluttyp <- NULL
ols <- ols[ols$birthdate >= 1841 & ols$birthdate < 1851, ]
ols <- age.window(ols, c(60, 100))
ids <- ols$id[ols$enter < 60.01]
ols <- ols[ols$id %in% ids, ]
indx <- tapply(ols$id, ols$id)
out <- tapply(ols$exit, ols$id, max)
ols$out <- out[indx]
ols <- ols[order(ols$id, -ols$exit, -ols$event),]
##sum(ols$event)
ols <- ols[!duplicated(ols$id), ]
##sum(ols$event)
ols$exit <- ols$out
ols$out <- NULL
summary(ols)
```

@@

In the following, we are interested in the *sex difference* in expected life after 60 (just as an illustration). The categorical variable `region` is included as an example of a confounder.

## Standard survival analyses

### Ordinary *Cox regression*

@@

### AFT regression

@@

@@

@@

## Sampling deaths

@@

### Ordinary *Cox regression*

```{r standard}
fit <- coxreg(Surv(enter, exit, event) ~ sex + region, data = sols)
summary(fit)
```

### AFT regression

@@

## OLS regression

>>>>

---
title: "Sampling deaths and OLS"
author: "Göran Broström"
date: "January 19, 2018"
output:
  bookdown::html_document2:
    number_sections: yes
    toc: yes
    toc_depth: 2
    toc_float: yes
  bookdown::pdf_document2:
    citation_package: natbib
    number_sections: yes
    toc: no
    toc_depth: 2
    keep_tex: true

---

# Introduction

@@

# Data

@@

```{r readalldata}
##source("R/get.file.R")
##skume <- get.file("skume")
## Not needed(?):
##vb0Tab <- skume
##vb0Tab$civst <- relevel(vb0Tab$civst, ref = "married")
##vb0Tab <- aggregate(vb0Tab[, c("event", "exposure")],
##                    vb0Tab[, c("sex", "age", "period", "cohort", "civst", "hisclass", "urban")],
##                    FUN = sum)
## End: Not needed(?)
ols <- obs[, c("id", "sex", "birthdate", "region", "civst", "sluttyp", "enter", "exit")]
ols$event <- ols$sluttyp == 2
ols$sluttyp <- NULL
ols <- ols[ols$birthdate >= 1841 & ols$birthdate < 1851, ]
ols <- age.window(ols, c(60, 100))
ids <- ols$id[ols$enter < 60.01]
ols <- ols[ols$id %in% ids, ]
indx <- tapply(ols$id, ols$id)
out <- tapply(ols$exit, ols$id, max)
ols$out <- out[indx]
ols <- ols[order(ols$id, -ols$exit, -ols$event),]
##sum(ols$event)
ols <- ols[!duplicated(ols$id), ]
##sum(ols$event)
ols$exit <- ols$out
ols$out <- NULL
summary(ols)
```

@@

In the following, we are interested in the *sex difference* in expected remaining life after 60 (*e60*) (just as an illustration). The categorical variable `region` is included as an example of a confounder.

# Standard survival analyses

## Ordinary *Cox regression*

@@

We cannot easily calculate *e60* from this figure (or table above), but the estimated median remaining life at 60 (*m60*) is easy to illustrate graphically. For that porpose we need to plot the estimated *survival functions*.
```{r }
plot(survfit(fit), col = 1:4, lty = 1:4, xlab = "Years after 60.", 
     ylab = "Remaining fraction", xlim = c(0, 30), axes = FALSE)
axis(1, at = c(0, 9, 13, 20, 30))
axis(2, at = c(0, 0.5, 1), las = 1)
box()
legend("bottomleft", legend = levels(strata(ols$region, ols$sex)), col = 1:4, lty = 1:4)
abline(h = 0)
abline(h = 0.5, lty = 2)
```

We can also make an exact calculation, which results in

```{r exactki}
x <- survfit(fit)
x
```

## AFT regression

@@

The *e60* for men in Skellefteå is 9.69, and for Skellefteå women it is $1.329 \times 9.69 = 12.88$, a difference of 3.19 years.

@@

The *e60* for men in Skellefteå is now 11.2, and for Skellefteå women it is $1.075 \times 11.2 = 12.04$, a difference of only 0.84 years.

@@

A lot more reasonable.

# Sampling deaths

@@

## Ordinary *Cox regression*

```{r standard}
fit <- coxph(Surv(enter, exit, event) ~ sex + region, data = sols)
survfit(fit)
```

## AFT regression

@@

# OLS regression
